---
title: "complicated_regression"
author: "Isabel Gomez"
date: "12/7/2020"
output: html_document
---

#loading packages and reading csv
```{r}
library(tidyverse)
library(rstanarm)
library(bayesplot)
library(bayesrules)
library(rstan)

#loading csv need to put your own pathway in
birth_data1 <-read_csv("birth_data_10,000.csv") #csv is too large to put on github
```

```{r reading- data, include=FALSE}
#selecting variables of interest
birth_data <- birth_data1 %>%
  select("mbrace", "pay", "no_mmorb", "mager9", "dmar")
```

#Exploring the data
```{r}
#count of unknown insurance
birth_data %>%
  filter(pay == 8) %>%
  count()

#count of unknown morbidity and insurance
birth_data %>%
  filter(pay == 8 | pay == 9 |no_mmorb == 9) %>%
  count()

#count of unknown morbidity
birth_data %>%
  filter(no_mmorb == 9) %>%
  count()
```


#Data Wrangling for white vs non-white and insurance vs no insurance
```{r binary-data-wrangling}
#Creating data frame with columns of interest
birth_binary <- birth_data %>%
  select(mbrace, pay, no_mmorb, mager9, dmar) %>%
  filter(no_mmorb != 9) %>%
  filter(pay != 9) %>% 
  filter(pay != 8) %>% #filtering out unknowns and others 
  drop_na() %>%
  mutate(
    new_payment =
     ifelse(pay %in% c(1, 2, 4, 5, 6), 1, 0) #changing insurance into insur vs no insur
         ) %>% 
   mutate(new_race = 
            ifelse(mbrace %in% c(2, 3, 4), 1, 0)) #changing race into non-white vs white

#Rewriting data for type of insurance and race from numbers to words. 
birth_binary  <- birth_binary %>%
  mutate(insurance = recode(new_payment, `1` = "insurance", `0` = 'AAno insurance'), 
           race = recode(new_race, `1` = "Non-white", `0` = "AAWhite"))

#changing insurance and race into factors 
morbidity_binary_data <-  birth_binary %>% 
  mutate(insurance = as.factor(birth_binary$insurance),
         race = as.factor(birth_binary$race),
         age = as.factor(birth_binary$mager9),
         marital = as.factor(birth_binary$dmar))
```

#Data wrangling for model with all racial categories
```{r race-data-wrangling}

#Creating data frame with columns of interest
race_data <- birth_data %>%
  filter(no_mmorb != 9) %>%
  filter(pay != 9) %>% 
  filter(pay != 8) %>% 
  mutate(new_payment = if_else(pay %in% c(1, 2, 4, 5, 6), 1,0)
    ) %>%
  drop_na() 
#Rewriting data for type of insurance and race from numbers to words. 
morbidity_data <-race_data %>%
  mutate(insurance = recode(new_payment, `1` = "insurance", `0` = "AA no insurance"), 
           race = recode(mbrace, `1` = "AAWhite", `2` = "Black", `3` = "American Indians and Alaska Natives", `4` = "Asian or Pacific Islander"))

#changing insurance and race into factors 
morbidity_data <- morbidity_data %>%
   mutate(insurance = as.factor(morbidity_data$insurance),
          race = as.factor(morbidity_data$race),
          age = as.factor(morbidity_data$mager9),
          marital = as.factor(morbidity_data$dmar))  
```

#Simple Model

First we fit our most simple model, which includes the binary variables of health insurance and race, adjusting for age and marital status.
```{r logistic-regression-simple-model}
set.seed(84735)
morbidity_simple_model <- stan_glm(
  no_mmorb ~ race + insurance + age + marital, 
  data = morbidity_binary_data, family = binomial,
  chains = 4, iter = 5000*2)
```

Next, we wanted to summarize our model.
```{r simple-model-summary}
simple_model_summary <- summary(morbidity_simple_model)
head(as.data.frame(simple_model_summary),-2)
```

```{r simple-model-odds}
#means
exp(0.36729956) #race
exp(-0.71808914) #insurance
```

Adjusting for age, marital status, and insurance, the odds that a non-white person experiences maternal morbidity is an average of 1.44 times the odds that a white person experiences morbidity.

Adjusting for age, marital status, and race, the odds that someone with insurance experiences maternal morbidity is an average of 0.487 times the odds that someone without insurance experiences morbidity.

```{r simple-model-intervals}
#credible intervals
exp(c(0.06499287	, 0.67682544)) #race
exp(c(-1.55097967, 0.04133771)) #insurance
```

Adjusting for age, marital status, and insurance, there is an 80% posterior probability that the odds that a non-white person experiences maternal morbidity is between 1.06 and 1.96 times the odds that a white person experiences morbidity.

Adjusting for age, marital status, and insurance, there is an 80% posterior probability that the odds that someone who has insurance experiences maternal morbidity is between 0.21 and 1.04 times the odds that someone who does not have insurance experiences morbidity.

```{r simple-model-dens}
#density and trace plots of simple model
mcmc_dens(morbidity_simple_model)
mcmc_trace(morbidity_simple_model)
```

#Complicated Model

We wil not interpret the results but just use this to perform comparison tests to see if this is  better than the simple model. If it's better then we could interpret. 

Slightly more complicated model with interaction term between binary variables of race and insurance. 
```{r complicated-model}
set.seed(84735)
complicated_morbidity_model <- stan_glm(
  no_mmorb ~ race + insurance + age + marital + race*insurance, 
  data = morbidity_binary_data, family = binomial,
  chains = 4, iter = 5000*2)
```

Summary of interaction model. 
```{r complicated-model-summary}
complicated_model_summary <- summary(complicated_morbidity_model)
head(as.data.frame(complicated_model_summary),-2)
```

```{r complicated-model-dens}
#density and trace plots of complicated model
mcmc_dens(complicated_morbidity_model)
mcmc_trace(complicated_morbidity_model)
```



# Race Model
```{r logistic-regression-model}
set.seed(84735)
race_morbidity_model <- stan_glm(
  no_mmorb ~ race + insurance + age + marital, 
  data = morbidity_data, family = binomial,
  chains = 4, iter = 5000*2)
```

Next, we wanted to summarize our model.
```{r race-model-summary}
race_model_summary <- summary(race_morbidity_model)
head(as.data.frame(race_model_summary),-2)
```

```{r odds-of-mean}
#mean
exp(c(-0.49408245, 0.06471849 , 0.68447106)) #native american, asian, black
exp(c(-0.72164298)) #insurance
```

Adjusting for age, marital status, and insurance, the odds that someone who is Native American experiences maternal morbidity is an average of 0.69 times the odds that someone who is white experiences morbidity.

Adjusting for age, marital status, and insurance, the odds that someone who is Asian or Pacific Islander experiences maternal morbidity is an average of 1.06 times the odds that someone who is white experiences morbidity.

Adjusting for age, marital status, and insurance, the odds that someone who is 
Black experiences maternal morbidity is an average of 1.98 times the odds that someone who is white experiences morbidity.

Adjusting for age, marital status, and race, the odds that someone with insurance experiences maternal morbidity is an average of 0.485 times the odds that someone without insurance experiences morbidity.

```{r odds-of-interval}
#credible intervals
exp(c(-1.2704155, 0.35641770)) #native american
exp(c(-0.4434041, 0.59677159)) #asian or pacific islander
exp(c(0.2830783, 1.09873140)) #black
exp(c(-1.5612151, 0.03870364)) #insurance
```

Adjusting for age, marital status, and insurance, there is an 80% posterior probability that the odds that a Native American experiences maternal morbidity is between 0.28 and 1.42 times the odds that a white person experiences morbidity.

Adjusting for age, marital status, and insurance, there is an 80% posterior probability that the odds that an Asian person or Pacific Islander experiences maternal morbidity is between 0.64 and 1.81 times the odds that someone who is white experiences morbidity.

Adjusting for age, marital status, and insurance, there is an 80% posterior probability that the odds that a Black person experiences maternal morbidity is between 1.32 and 3 times the odds that a white person experiences morbidity.

Adjusting for age, marital status, and race, there is an 80% posterior probability that the odds that someone with insurance experiences maternal morbidity is between 0.21 and 1.03 times the odds that someone without insurance experiences morbidity.


```{r race-model-density}
#density and trace plots of race model
mcmc_dens(race_morbidity_model)
mcmc_trace(race_morbidity_model)
```
