---
title: "Building Logistic Regression"
author: "Isabel Gomez"
date: "12/6/2020"
output: html_document
---

#loading packages
```{r}
library(tidyverse)
library(rstanarm)
library(bayesplot)
library(bayesrules)
library(rstan)
```

Reading in data
```{r}
birth_data1 <- read.csv("birth_data1.csv")

```


```{r}
#Rewriting data for type of insurance and race  from numbers to words. 
birth_data <- birth_data1 %>%
  mutate(insurance = recode(new_payment, `1` = "insurance", `0` = 'AAno insurance'), 
           race = recode(new_race, `1` = "Non-white", `0` = "AAWhite"))

#changing insurance and race into factors 
morbidity_data <-  birth_data %>% mutate(insurance = as.factor(birth_data$insurance),
                                race = as.factor(birth_data$race)) 
```



First we fit our most simple model, which includes health insurance and race, adjusting for age and marital status.
```{r logistic-regression-model}
set.seed(84735)
morbidity_model <- stan_glm(
  no_mmorb ~ race + insurance + mager9 + dmar, 
  data = morbidity_data, family = binomial,
  chains = 4, iter = 5000*2)
```

Next, we wanted to summarize our model.
```{r}
model_summary <- summary(morbidity_model)
head(as.data.frame(model_summary),-2)
```

```{r}
#means
exp(0.39084452) #race
exp(-0.66459581) #insurance
```

Being non-white increases the typical odds of maternal morbidity by an average of 1.47.

Having insurance decreases the typical odds of maternal morbidity by an average of 0.5144.

```{r}
#credible intervals
exp(c(0.08348851, 0.70422622)) #race
exp(c(-1.36319390, -0.01304253)) #insurance
```

For a non-white person, there is an 80% posterior chance that the typical increase in the odds of maternal morbidity is between 1.08 and 2.02.

For a person with insurance, there is an 80% posterior chance that the typical decrease in the odds of maternal morbidity is between 0.25 and 0.98.

```{r}
#density and trace plots of model 1
mcmc_dens(morbidity_model)
mcmc_trace(morbidity_model)
```

Slightly more complicated model with interaction term between race and insurance. 
```{r}
set.seed(84735)
morbidity_model_2 <- stan_glm(
  no_mmorb ~ race + insurance + mager9 + dmar + race*insurance, 
  data = morbidity_data, family = binomial,
  chains = 4, iter = 5000*2)
```

Summary of interaction model. 
```{r}
model_summary2 <- summary(morbidity_model_2)
head(as.data.frame(model_summary2),-2)
```

```{r}
#density and trace plots of model 2
mcmc_dens(morbidity_model_2)
mcmc_trace(morbidity_model_2)
```





