---
title: "complicated_regression"
author: "Isabel Gomez"
date: "12/7/2020"
output: html_document
---

#loading packages and reading csv
```{r}
library(tidyverse)
library(rstanarm)
library(bayesplot)
library(bayesrules)
library(rstan)

#loading csv need to put your own pathway in
birth_data1 <-read_csv("birth_data_10,000.csv") #csv is too large to put on github
```

```{r reading- data, include=FALSE}
#selecting variables of interest
birth_data <- birth_data1 %>%
  select("mbrace", "pay", "no_mmorb", "mager9", "dmar")
```

#Exploring the data
```{r}
#count of unknown insurance
birth_data %>%
  filter(pay == 8) %>%
  count()

#count of unknown morbidity and insurance
birth_data %>%
  filter(pay == 8 | pay == 9 |no_mmorb == 9) %>%
  count()

#count of unknown morbidity
birth_data %>%
  filter(no_mmorb == 9) %>%
  count()
```


#Data Wrangling for white vs non-white and insurance vs no insurance
```{r binary-data-wrangling}
#Creating data frame with columns of interest
birth_binary <- birth_data %>%
  select(mbrace, pay, no_mmorb, mager9, dmar) %>%
  filter(no_mmorb != 9 | pay != 8 | pay !=9) %>% #filtering out unknowns and others
  mutate(
    new_payment =
     ifelse(pay %in% c(1, 2, 4, 5, 6), 1, 0) #changing insurance into insur vs no insur
         ) %>% 
   mutate(new_race = 
            ifelse(mbrace %in% c(2, 3, 4), 1, 0)) #changing race into non-white vs white

#Rewriting data for type of insurance and race from numbers to words. 
birth_binary  <- birth_binary %>%
  mutate(insurance = recode(new_payment, `1` = "insurance", `0` = 'AAno insurance'), 
           race = recode(new_race, `1` = "Non-white", `0` = "AAWhite"))

#changing insurance and race into factors 
morbidity_binary_data <-  birth_binary %>% mutate(insurance =
                                                    as.factor(birth_data2$insurance),
                                                  race = as.factor(birth_data2$race))
```

#Data wrangling for model with all racial categories
```{r race-data-wrangling}

#Creating data frame with columns of interest
race_data <- birth_data %>%
  filter(no_mmorb != 9 | pay !=8 | pay != 9 ) %>% 
  mutate(new_payment = if_else(pay %in% c(1, 2, 4, 5, 6), 1,0)
    )
#Rewriting data for type of insurance and race from numbers to words. 
morbidity_data <-race_data %>%
  mutate(insurance = recode(new_payment, `1` = "insurance", `0` = "AA no insurance"), 
           race = recode(mbrace, `1` = "AAWhite", `2` = "Black", `3` = "American Indians
                         and Alaska Natives", `4` = "Asian or Pacific Islander"))

#changing insurance and race into factors 
morbidity_data <- morbidity_data %>%
          mutate(insurance = as.factor(birth_data1$insurance ),
                                race = as.factor(birth_data1$race)) 
```

#Simple Model

First we fit our most simple model, which includes health insurance and race, adjusting for age and marital status.
```{r logistic-regression-simple-model}
set.seed(84735)
morbidity_simple_model <- stan_glm(
  no_mmorb ~ race + insurance + mager9 + dmar, 
  data = morbidity_data, family = binomial,
  chains = 4, iter = 5000*2)
```

Next, we wanted to summarize our model.
```{r simple-model-summary}
simple_model_summary <- summary(morbidity_simple_model)
head(as.data.frame(simple_model_summary),-2)
```

```{r simple-model-odds}
#means
exp(0.39084452) #race
exp(-0.66459581) #insurance
```

Being non-white increases the typical odds of maternal morbidity by an average of 1.47.

Having insurance decreases the typical odds of maternal morbidity by an average of 0.5144.

```{r simple-model-intervals}
#credible intervals
exp(c(0.08348851, 0.70422622)) #race
exp(c(-1.36319390, -0.01304253)) #insurance
```

For a non-white person, there is an 80% posterior chance that the typical increase in the odds of maternal morbidity is between 1.08 and 2.02.

For a person with insurance, there is an 80% posterior chance that the typical decrease in the odds of maternal morbidity is between 0.25 and 0.98.

```{r simple-model-dens}
#density and trace plots of simple model
mcmc_dens(morbidity_simple_model)
mcmc_trace(morbidity_simple_model)
```

#Complicated Model

We wil not interpret the results but just use this to perform comparison tests to see if this is  better than the simple model. If it's better then we could interpret. 

Slightly more complicated model with interaction term between race and insurance. 
```{r complicated-model}
set.seed(84735)
complicated_morbidity_model <-  no_mmorb ~ race + insurance + mager9 + dmar +
  race*insurance, 
  data = morbidity_data, family = binomial,
  chains = 4, iter = 5000*2)
```

Summary of interaction model. 
```{r complicated-model-summary}
complicated_model_summary <- summary(complicated_morbidity_model)
head(as.data.frame(complicated_model_summary),-2)
```

```{r complicated-model-dens}
#density and trace plots of complicated model
mcmc_dens(complicated_morbidity_model)
mcmc_trace(complicated_morbidity_model)
```



# Race Model
```{r logistic-regression-model}
set.seed(84735)
race_morbidity_model <- stan_glm(
  no_mmorb ~ race + insurance + mager9 + dmar, 
  data = morbidity_data, family = binomial,
  chains = 4, iter = 5000*2)
```

Next, we wanted to summarize our model.
```{r race-model-summary}
race_model_summary <- summary(race_morbidity_model)
head(as.data.frame(race_model_summary),-2)
```

```{r odds-of-mean}
#mean
exp(c(-0.5362581 ,0.0613946 , 0.7050168)) #native american, asian, black
exp(c(-0.7248291)) #insurance
```
Being Native American decreases the typical odds of maternal morbidity by an average of 0.59.
Being Asian or Pacific Islander increases the typical odds of maternal morbidity by an average of 1.06.
Being Black increases the the typical odds of maternal morbidity by an average of 2.02.
Having insurances decreases the typical of maternal morbidity by an average of 0.48. 

```{r odds-of-interval}
#credible intervals
exp(c(-1.30355734, 0.30452888	)) #native american
exp(c(-0.44791471, -0.60235732)) #asian or pacific islander
exp(c(0.30438839,1.11749202)) #black
exp(c(-1.55846819, 0.03313465)) #insurance
```

For a Native Ameircan person, there is an 80% posterior chance that the typical increase in the odds of maternal morbidity is between 0.27 and 1.36.

For an Asian person, there is an 80% posterior chance that the typical increase in the odds of maternal morbidity is between 0.64 and 0.55.

For a Black person, there iss an 80% posterior chance that the typical increase in the odds of maternal morbidity it between 1.36 and 3.06. 

For a person with insurance, there is an 80% posterior chance that the typical decrease in the odds of maternal morbidity is between  0.21 and 1.03.


```{r race-model-density}
#density and trace plots of race model
mcmc_dens(race_morbidity_model)
mcmc_trace(race_morbidity_model)
```
