---
title: "pregnancy_risk_exploration"
author: "Elisabeth Nesmith"
date: "12/12/2020"
output: html_document
---

```{r}
library(tidyverse)
library(rstanarm)
library(bayesplot)
library(bayesrules)
library(rstan)

#loading csv need to put your own pathway in
birth_data1 <-read_csv("birth_data_10,000.csv")
```

#Selecting variables of interest 
```{r}
pregnancy_data <- birth_data1 %>%
  select("mbrace", "pay", "rf_pdiab", "rf_gdiab", "rf_phype", "rf_ghype", "rf_ehype", "rf_ppterm", "mager9", "dmar")

#filter unknowns and make pregnancy risk one variable
pregnancy_data1 <- pregnancy_data %>% 
  filter(rf_pdiab != "U") %>% 
  filter(rf_gdiab != "U") %>%
  filter(rf_phype != "U") %>% 
  filter(rf_ghype != "U") %>%
  filter(rf_ehype != "U") %>%
  filter(rf_ppterm != "U") %>% 
  mutate(pregnancy_risk = 
           ifelse(rf_pdiab == "Y", 1, 
           ifelse(rf_gdiab == "Y", 1,
           ifelse(rf_phype == "Y", 1,
           ifelse(rf_ghype == "Y", 1,
           ifelse(rf_ehype == "Y", 1,
           ifelse(rf_ppterm == "Y", 1, 0)))))))
  
```


```{r}
#check counts of people who are at increased risk
pregnancy_data1 %>%
  select(pregnancy_risk) %>%
  group_by(pregnancy_risk) %>%
  summarize(count = n())
```

#Data Wrangling for white vs non-white and insurance vs no insurance
```{r}
#Creating data frame with columns of interest
pregnancy_binary <- pregnancy_data1 %>%
  select(mbrace, pay, pregnancy_risk, mager9, dmar) %>%
  filter(pay != 9) %>% 
  filter(pay != 8) %>% #filtering out unknowns and others 
  drop_na() %>%
  mutate(
    new_payment =
     ifelse(pay %in% c(1, 2, 4, 5, 6), 1, 0) #changing insurance into insur vs no insurance
         ) %>% 
   mutate(new_race = 
            ifelse(mbrace %in% c(2, 3, 4), 1, 0)) #changing race into non-white vs white

#Rewriting data for type of insurance and race from numbers to words. 
pregnancy_binary  <- pregnancy_binary %>%
  mutate(insurance = recode(new_payment, `1` = "insurance", `0` = 'AAno insurance'), 
           race = recode(new_race, `1` = "Non-white", `0` = "AAWhite"))

#changing insurance, age, marital status, and race into factors 
pregnancy_binary_data <- pregnancy_binary %>% 
  mutate(insurance = as.factor(pregnancy_binary$insurance),
         race = as.factor(pregnancy_binary$race),
         age = as.factor(pregnancy_binary$mager9),
         marital = as.factor(pregnancy_binary$dmar))
```

#Data wrangling for model with all racial categories
```{r race-data-wrangling}

#Creating data frame with columns of interest
race_data <- pregnancy_data1 %>%
  select(mbrace, pay, pregnancy_risk, mager9, dmar) %>%
  filter(pay != 9) %>% 
  filter(pay != 8) %>% 
  mutate(new_payment = if_else(pay %in% c(1, 2, 4, 5, 6), 1,0)
    ) %>%
  drop_na() 

#Rewriting data for type of insurance and race from numbers to words. 
preg_risk_data <-race_data %>%
  mutate(insurance = recode(new_payment, `1` = "insurance", `0` = "AA no insurance"), 
           race = recode(mbrace, `1` = "AAWhite", `2` = "Black", `3` = "American Indians and Alaska Natives", `4` = "Asian or Pacific Islander"))

#changing insurance and race into factors 
preg_risk_data <- preg_risk_data %>%
   mutate(insurance = as.factor(preg_risk_data$insurance),
          race = as.factor(preg_risk_data$race),
          age = as.factor(preg_risk_data$mager9),
          marital = as.factor(preg_risk_data$dmar))
```

#Descriptive Statistics
```{r}
# pregnancy risk counts, 0 means no sign of risk, 1 means risk
pregnancy_binary_data %>%
  select(pregnancy_risk) %>%
  group_by(pregnancy_risk) %>%
  summarize(count = n())

#risk by binary race
pregnancy_binary_data %>%
  select(pregnancy_risk, race) %>%
  filter(pregnancy_risk == 1) %>% 
  group_by(race) %>%
  summarize(count = n())

#risk by categorical race
preg_risk_data %>%
  select(pregnancy_risk, race) %>%
  filter(pregnancy_risk == 1) %>% 
  group_by(race) %>%
  summarize(count = n())

#risk by insurance
pregnancy_binary_data %>%
  select(pregnancy_risk, insurance) %>%
  filter(pregnancy_risk == 1) %>% 
  group_by(insurance) %>%
  summarize(count = n())
```

```{r}
# payment counts
pregnancy_binary_data %>%
  select(insurance) %>%
  group_by(insurance) %>%
  summarise(pay_count = n())
```
```{r}
# binary race counts
pregnancy_binary_data %>%
  select(race) %>%
  group_by(race) %>%
  summarise(race_count = n())
```
```{r}
# categorical race counts
preg_risk_data %>%
  select(race) %>%
  group_by(race) %>%
  summarise(race_count = n())
```


#Simple Model

First we fit our most simple model, which includes the binary variables of health insurance and race, adjusting for age and marital status.

```{r logistic-regression-simple-model}
set.seed(84735)
risk_simple_model <- stan_glm(
  pregnancy_risk ~ race + insurance + age + marital, 
  data = pregnancy_binary_data, family = binomial,
  chains = 4, iter = 5000*2)
```

Next, we wanted to summarize our model.
```{r simple-model-summary}
simple_model_summary <- summary(risk_simple_model)
head(as.data.frame(simple_model_summary),-2)
```

```{r simple-model-odds}
#means
exp(0.2664565) #race
exp(0.6201262) #insurance
```

Adjusting for age, marital status, and insurance, the odds that a non-white person has increased risk during pregnancy is an average of 1.3 times the odds that a white person has increased risk.

Adjusting for age, marital status, and race, the odds that someone with insurance has increased risk during pregnancy is an average of 1.86 times the odds that someone without insurance has increased risk.

```{r simple-model-intervals}
#credible intervals
exp(c(0.1847, 0.3475453)) #race
exp(c(0.4112, 0.8310631)) #insurance
```

Adjusting for age, marital status, and insurance, there is an 80% posterior probability that the odds that a non-white person has increased risk during pregnancy is between 1.2 and 1.4 times the odds that a white person has increased risk.

Adjusting for age, marital status, and race, there is an 80% posterior probability that the odds that someone who has insurance has increased risk during pregnancy is between 1.51 and 2.29 times the odds that someone who does not have insurance has increased risk.

```{r simple-model-dens}
#density and trace plots of simple model
mcmc_dens(risk_simple_model)
mcmc_trace(risk_simple_model)
```

#Complicated Model

We will not interpret the results but just use this to perform comparison tests to see if this is  better than the simple model. If it's better then we could interpret. 

Slightly more complicated model with interaction term between binary variables of race and insurance. 
```{r complicated-model}
set.seed(84735)
complicated_risk_model <- stan_glm(
  pregnancy_risk ~ race + insurance + age + marital + race*insurance, 
  data = pregnancy_binary_data, family = binomial,
  chains = 4, iter = 5000*2)
```

Summary of interaction model. 
```{r complicated-model-summary}
complicated_model_summary <- summary(complicated_risk_model)
head(as.data.frame(complicated_model_summary),-2)
```

```{r complicated-model-dens}
#density and trace plots of complicated model
mcmc_dens(complicated_risk_model)
mcmc_trace(complicated_risk_model)
```

# Race Model
```{r logistic-regression-model}
set.seed(84735)
race_risk_model <- stan_glm(
  pregnancy_risk ~ race + insurance + age + marital, 
  data = preg_risk_data, family = binomial,
  chains = 4, iter = 5000*2)
```

Next, we wanted to summarize our model.
```{r race-model-summary}
race_model_summary <- summary(race_risk_model)
head(as.data.frame(race_model_summary),-2)
```

```{r odds-of-mean}
#mean
exp(c(0.2180731, 0.2351377, 0.2788102)) #native american, asian, black
exp(c(0.6215643)) #insurance
```

Adjusting for age, marital status, and insurance, the odds that someone who is Native American has increased risk during pregnancy is an average of 1.24 times the odds that someone who is white has increased risk.

Adjusting for age, marital status, and insurance, the odds that someone who is Asian or Pacific Islander has increased risk during pregnancy is an average of 1.26 times the odds that someone who is white has increased risk.

Adjusting for age, marital status, and insurance, the odds that someone who is 
Black has increased risk during pregnancy is an average of 1.32 times the odds that someone who is white has increased risk.

Adjusting for age, marital status, and race, the odds that someone with insurance experiences increased risk during pregnancy is an average 1.86 times the odds that someone without insurance has increased risk.

```{r odds-of-interval}
#credible intervals
exp(c(-0.10750214, 0.5343003)) #native american
exp(c(0.09083989, 0.3781204)) #asian or pacific islander
exp(c(0.18496426, 0.3727168)) #black
exp(c(0.41427921, 0.8367461)) #insurance
```

Adjusting for age, marital status, and insurance, there is an 80% posterior probability that the odds that a Native American has increased risk during pregnancy is between 0.89 and 1.71 times the odds that a white person has increased risk.

Adjusting for age, marital status, and insurance, there is an 80% posterior probability that the odds that an Asian person or Pacific Islander has increased risk during pregnancy is between 1.09 and 1.46 times the odds that someone who is white has increased risk.

Adjusting for age, marital status, and insurance, there is an 80% posterior probability that the odds that a Black person has increased risk during pregnancy is between 1.2 and 1.45 times the odds that a white person has increased risk.

Adjusting for age, marital status, and race, there is an 80% posterior probability that the odds that someone with insurance has increased risk during pregnancy is between 1.51 and 2.3 times the odds that someone without insurance has increased risk.

```{r race-model-density}
#density and trace plots of race model
mcmc_dens(race_risk_model)
mcmc_trace(race_risk_model)
```

#comparison of models using ELPD
```{r}
loo_1 <- loo(risk_simple_model)
loo_2 <- loo(complicated_risk_model)
loo_3 <- loo(race_risk_model)
loo_compare(loo_1, loo_2, loo_3)
```

